Fitness effects of mutations to SARS-CoV-2 proteins

### 2023-01-15 paper/paper.tex
third Intro paragraph

An alternative is sequence-based estimates, which so far have focused largely on clade growth and identification of mutations that mediate immune escape or increase transmissibility \citep{obermeyer2022analysis,lee2022inferring}.
Mutations that are repeatedly observed at the base of clades that grow faster than the background likely contribute to growth.
Such adaptive mutations are only a small minority of all available mutations and which mutations are adaptive depends on the environment (immunity) and the genetic background.
The great majority of mutations, however, have either small fitness effects (approximately neutral) or are deleterious to viral replication.
Here, we describe an approach to quantify the fitness cost of these mutations based on comparing expected versus observed counts.

\rancomment{older paper on flu where not counts but trajectories were used to partition sequences into good and bad \citep{strelkowa2021clonal}}

>>>>

An alternative to experiments is to estimate the effects of mutations by analyzing naturally occurring viral sequences.
The amount of data available for such analyses has increased dramatically over the last few years with the full-genome sequencing of SARS-CoV-2 viruses from millions of human infections~\citep{?}.
So analyses of these sequences have focused largely on analyzing expanding viral clades to identify mutations that mediate immune escape or increase transmissibility~\citep{obermeyer2022analysis,lee2022inferring}.
The basic idea of these prior analyses is that mutations that repeatedly appear in clades that subsequently expand in relative frequency are likely beneficial to the virus.
However, only a small minority of all possible mutations are beneficial, with the great majority being either nearly neutral or deleterious to viral fitness.
For purposes such as identifying functionally constrained drug targets, it is important to accurately estimate the effects of neutral or deleterious mutations as well as beneficial ones.
\jdbcomment{Richard, I could not quite figure out how to fit in \citep{strelkowa2021clonal}, let me know if you have ideas.}
